A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-20 manuscript.Rmd
Second round of comments; more after dinner

<!-- define replicability and then... 
 define reproducibility -->
Even worse, researchers might often not even be able to achieve reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models.
With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.

In fact, the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- examples missing -->
However, ensuring reproducibility of computational analyses is still tedious and difficult, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
For example, @hardwicke2018 attempted to replicate open materials in the journal Cognition.
Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced —-- even with the original authors' assistance.
@obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.

From our personal experience, the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step.
If reproduction would require merely minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
<!-- Guiding principle: spending machine compute time instead of human research time -->
To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.
To this end, we demand that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows.

<!-- Which problems are solved by which tool (like in Andreas talk)-->
<!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->
We identified the following common four causes for non-reproducibility:

1.	Multiple inconsistent versions of code and data; for example, the dataset has changed over time because outliers were removed at a later stage, or the code has changed because a bug was removed at some point in time;
2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.

We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:

@@

This tutorial follows the recommendations of @Peikert2019 and uses _Git_ for version control, _R Markdown_ for dynamic document creation, _Make_ for dependency tracking, and _Docker_ for software management.
The interplay of these four components is shown in Figure \@ref(fig:schematic).
However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].
At several points, we suggest other viable alternatives.

@@

This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8. which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact using a unique DOI in a persistent location.

@@

Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.
However, not all steps are necessary to achieve reproducibility.
We have marked those steps that can be omitted as "optional" in the section headings.

@@

This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:

@@

```r
# -- type this on the R console --
if(!requireNamespace("remotes"))install.packages("remotes")
remotes::install_github("aaronpeikert/repro")
library("repro")
```

The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with the suggested tools for reproducibility.
For example, the package can check if your computer has all the required software packages installed and ready.

@@

If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
Please make sure that all three commands pass before you continue with the tutorial.
This process is only required once per computer and does not need to be repeated for each project.

# Planning Phase

@@

Before incorporating changes from someone else, they can be reviewed and adapted.
This review step is particularly useful for remote collaboration, because collaborators can be explicitly pointed to changes and asked to approve them.
Therefore, it is sometimes useful to create a pull request even though you have permission to push, because it invites your collaborators to comment on, change, or approve proposed changes.

@@

This process can be executed via Github's web interface and can work well with collaborators who do not actually want to bother with installing and running the entire code base of a project.

# The Data Analysis Phase

So far, we have covered the generation of dynamic documents with [Markdown] and [R Markdown], and tracking changes with [Git] and [GitHub]. These are two of the four components [see @Peikert2019] to achieve long-term reproducibility. To complete computational reroducibility, researchers need to properly handle dependency relations between their files (internal dependencies) and all dependencies on external software, most importantly dependend R packages (external dependencies). In the following, we will show how the `repro` package supports researchers in mapping all dependencies completely and unambiguously.

@@

A reader of a reproducible research project should be able to easily derive how it can be reproduced (ideally by a statement in your [README]).
To do that, it is often crucial to understand how different project components relate to each other and depend on each other.
For example, you cannot run an analysis without data, or you cannot render a statistical report without having the results from the statistical analysis that is reported.
Dependency tracking means that you capture and map the interrelations between all components of your analysis.
The "components" of an analysis are typically all involved files including raw data files, intermediate files (such as a preprocessed data file), and final results (such as a rendered manuscript in PDF format).
With larger projects, there is typically a higher degree of modularization because code fragments are written for re-use in other projects (or are already re-used). Also, more complicated empirical research projects may have different hypotheses that are answered with different data sets and different statistical models, such that it often makes sense to split the different analyses into different files. Reproducibility then demands the careful documentation and mapping of all necessary steps in the correct order.

Make is a build automation tool for software projects that typically are challenged by maintaing and resolving a variety of dependencies in the correct order when compiling a piece of software-
We found _Make_ to be particularly useful for our purposes for two reasons.
First, it is language-agnostic, therefore enabling us to map steps not only taken within R but also across different programming languages or other external programs.
Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning. 
That is, Make determines automatically which branches in the dependency graph needs updating when files have changed and does not necessarily recompute or rebuild everything.

@@

If a target file already exists, Make only recreates it if necessary.
If a dependency is newer than its target, make recreates the target to reflect the updated dependency.
In our case, when we (re)build a manuscript and some dependent data file has changed, the data analysis is automatically rerun.
This avoids mistakes in the reporting that are due to reports and analyses being out-of-sync.

<!--
AP: I suggest to remove this simply because it distracts from our primary tool and we need to shorten
One alternative that is even smarter (in terms of automatically inferring dependencies) than Make is the R package [`targets` [@targets2021]](https://books.ropensci.org/targets/)[^targets].
`targets` can infer more about your project than Make, because it analyzes the R code you write to build the dependency tree.
For this inference to work well, you have to write everything in R, and you have to write everything strictly functional.

[^targets]: https://books.ropensci.org/targets/
-->

@@

Writing Makefiles manually is cumbersome for larger projects. 
Therefore, the `repro` package provides means to automatically write Makefiles for a given R project.
To this end, `repro` extends the YAML metadata header of R Markdown files. `repro` introduces a new first-level tag `repro:` with second-level tags (`scripts:`, `packages:`, and `data:`) that take lists of dependent files. For example, if an R Markdown document depends on a further file with R code, the header would look like this:

@@

To automatically generate the Makefile, `repro` provides a command that goes through all R Markdown files in the project, parses their YAML metadata, and infers a `Makefile` from them:

@@

The generated `Makefile` automatically contains separate targets for each R Markdown documents.
The names of the Makefile targets are identical to the Markdown file names.
For example, to generate the Markdown file `README.md` from the R Markdown `README.Rmd`, you can run:

@@

It is common practice to name the very first target in a Makefile "all" and have it include all important files of the project.
If you do not specify a target when calling Make, Make defaults to this first target.
`repro` reminds you of that common practice and suggests which file you might want to add to the target `all`.
Thus, if researchers who try to reproduce a given project and are unaware of its potentially various contents, will always build all reproducible artifacts of a repository when calling Make without further arguments:

@@

Running this command fully reproduces the entire project -- but only if you have all required software installed. Next, we will describe how `repro` helpts to manage all software dependencies automatically.

@@

Modern data analysis is only feasible because researchers rely on much external software. 
The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
External dependencies are unavoidable but, at the same time, a major threat to long-term reproducibility.
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. a software update or bug fixe, risks that balance.
At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Containers directly address such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
There are two major strategies helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they additionally need on top of these pre-made images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
Our image includes Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`

@@

From the `Dockerfile` you can then build an image.
While you can do so manually, we suggest to again use Make ( see [Internal Dependencies]).
The `Makefile` built by `repro` does contain a recipe for creating the `Dockerfile`.

@@

Importantly, the `Makefile` created by `repro` includes a mechanism to execute commands in the Docker container instead of locally.
This mechanism is triggered by the argument `DOCKER=TRUE` appended to any make call. 
The following call builds all targets inside the container instead of in the local computing environment:

```bash
# this should be run in the terminal
make DOCKER=TRUE
```

Last, to create or update the container and all Make targets, `repro` provides a single function `automate()` that calls `automate_docker()` and `automate_make()`.

You can also manually edit the `Dockerfile` by changing the file `.repro/Dockerfile_manual` to install software other than R packages.
Remember to run `repro::automate()` every time you made edits to `.repro/Dockerfile_manual`.
We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.

@@

Setting up and running Docker is not always painless, because it requires a complicated mental model of the underlying virtual computer and, specifically, knowledge about Linux environments.
An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
`renv` isolates all R packages used in a project and records there version as well as the source from where they where installed.
This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
`renv` therefore is already of great use for reproducibility [@vanlissa2020worcs].
However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS (Basic Linear Algebra Subprograms, responsible for carrying out many computations like matrix multiplication needed in the linear regression) system in use and so forth.
Most importantly the R version itself has to match.
For example, @Peikert2019 illustrate how the generation of random numbers across different R versions is a threat to reproducibility.

@@

In a public repository, you may want to mark important stages of your project, like preregistration, preprint, submission and acceptance at a scientific journal.
We suggest to archive these milestones individually.
A reproducible project is only then properly archived when it is stored at least at two independent storage locations and reproducible from all locations. <!--AP citation? -->
To really make sure that a given project is reproducible, one should aim for an explicit test of reproduction before the project is published. 
We recommend to check out and reproduce the project in a new directory in a "fresh" environment (in particular, a new R session should be used for this). 
For maximal validitiy of this test, we suggest to ask a coworker or collaborator to reproduce the project independently from you.
The sections [Creating a DOI] and [Creating a Release] describe how you can easily create three different openly accessible longterm archived copies of your project.

When you disseminate multiple versions of a document (e.g., at different stages of a review process), it seems good practice to specify the exact version of code and data used to create a given document (e.g. the final type-set version of a journal article).
To create a clear link between the repository version and a given end result you can automatically include the unique hash identifying the project state in the git version history.
<!--Either you use Git from within R:

```{r}
system2("git", c("rev-parse",  "--short",  "HEAD"), stdout = TRUE)
```
-->
To this end, we use the `gert` package:

@@

Typical open research project will have their contents available from multiple sites, e.g. from GitHub, preregistration platforms, preprint servers, or publisher websites. 
Under these circumstances it is important to provide unique links 
Digital Object Identifiers (DOI) tie all these locations together under one link [@derisiWhatWhysDOIs2003].
Not only can a DOI resolve to several locations, several parts of a research project can get their own DOI besides the DOI of the publication.
This is useful to link or cite a specific part of a research project, such as the preregistration, an illustration in the paper or the data that was gathered.
The DOI can also contain information about the license of its target.
<!-- public releases like preregistration, preprint -->
zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides DOIs for these archives.
While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
Please note that you can only link public GitHub repositories to zenodo.
You may log into zenodo.org through your GitHub account.

@@

After you have linked a GitHub repository, you trigger the archival by creating a GitHub release, as described in the next section.

@@

This image allows anyone with a computer equipped with Docker to reproduce the analysis (if the repository contains all code and data and has no other external dependencies).
Last, one can convert the Docker image to a full bootable system image.
Such a system image can subsequently be booted on any personal computer without relying on any software or service and therefore grantees reproducibility for decades to come.
Currently, `repro` does not yet facilitate this conversion, but we hope to include this feature soon.

>>>>

<!-- define replicability and then... 
 define reproducibility -->
Even worse, researchers might often not even be able to achieve reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models.
With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer. <!-- CJ this final claim is questionable, at least before introducing the concept of containerization. Just using data and procedures, you cannot guarantee reproducibility on any computer.-->

In fact, the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- examples missing -->
However, ensuring reproducibility of computational analyses is still tedious and difficult, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
For example, @hardwicke2018 attempted to replicate studies ostensibly adhering to the standards of open science in the journal Cognition.<!-- CJ How do you replicate open materials? At least in the social sciences "open materials" is taken to mean that you have used a permissive license on a newly developed questionnaire or other research material. -->
Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced —-- even with the original authors' assistance.
@obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.<!-- CJ It does not automatically follow that the problem you describe necessitates reproducible workflows. I would procede as follows: 1) use this example to conclude the existence of a problem (even "open" studies do not reproduce), 2) present workflows as a solution. -->

From our personal experience,<!--CJ This undermines the credibility of the argument. This is not an opinion piece--> the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step. <!-- CJ I am not sure this is the primary reason. Indeed, incentives are important - but in that context, perhaps reference DORA? I think you were just on a good track, arguing that even with the best intentions, researchers often fail to make their work reproducible. Your workflow will ensure that their next attempt is succesful.-->
If reproduction would require merely minutes instead of hours,<!-- CJ Now we're back to ease-of-use, not incentives.-->
it could greatly facilitate collaboration within and across research projects. <!-- CJ This is a different, third argument: In order to collaborate effectively, researchers should be looking at exactly the same data, code and results. I suggest to organize these arguments, and address them each exhaustively in turn. --> 
<!-- Guiding principle: spending machine compute time instead of human research time -->
To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.<!-- CJ I understand what you mean, but it's not yet communicated effectively because you have not yet made explicit that reproducible projects will indeed facilitate collaboration, and increase opportunities for collaboration (e.g., with strangers through github), and will therefore lead to a boost in productivity. Also - ask yourself if this boost in productivity is convincing to individual researchers, your target audience. It seems like this would be more convincing to management and policy makers.-->
To this end, we demand <!-- CJ It's you creating the product, so.. you demand this of yourself? Or do you mean that you demand this of your readership? Either way, strange choice of words.-->
that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows. <!-- CJ This is ambiguous. The point of "understandable by researchers and automatically reproducible by computers" requires clarification. "Minimal constraints to the workflow" is also not clear. Aren't you providing readers with a workflow? -->

<!-- Which problems are solved by which tool (like in Andreas talk)-->
<!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->
We identified the following common four causes for non-reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->

1.	Multiple inconsistent versions of code and data; for example, the dataset has changed over time because outliers were removed at a later stage, or the code has changed because a bug was removed at some point in time;
2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.
<!-- CJ 5. Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph. The opening paragraph should at least:
1) Introduce the topic area
2) Introduce the problem/knowledge gap
3) Introduce how you will fill this knowledge gap-->

@@

This tutorial follows the recommendations of @Peikert2019 and uses _Git_ for version control, _R Markdown_ for dynamic document creation, _Make_ for dependency tracking, and _Docker_ for software management. <!-- CJ None of these terms have been defined yet-->
The interplay of these four components is shown in Figure \@ref(fig:schematic).
However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].<!-- CJ But readers do not yet know what these concepts are, nor what tools exist. At this point in the writing, I think you need to take them by the hand and explain. After doing that, you can let them know that there are alternatives.-->
At several points, we suggest other viable alternatives.

@@

This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8. which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].<!-- CJ Is this really the case? Seems like there is little attention to the TOP guidelines, and the present paper is much more widely applicable. Maybe make the general point that workflows can help address formal guidelines for transparency, like TOP and FAIR - and consider citing WORCS for more information. Then explain that your workflow is flexible enough to go beyond--> 
We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.<!-- CJ Is this the description of the workflow?-->
We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact in a persistent location, and be assigned a DOI to increase findability.<!--CJ All DOIs are unique, and it is not stored using a DOI-->

@@

Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.<!-- CJ What do you mean here? -->
However, not all steps are necessary to achieve reproducibility.
We have marked those steps that can be omitted as "optional" in the section headings.

@@

This tutorial is targeted at scholars with at least intermediate R knowledge<!--CJ This sounds like something that belongs in the beginning of the introduction; it's not part of setup--> and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]: <!-- CJ This also seems like something that should be mentioned in the intro; once people are setting up their systems, they should havebeen told what they're setting them up for-->

@@

```r
# -- type this on the R console --
if(!requireNamespace("remotes"))install.packages("remotes")
remotes::install_github("aaronpeikert/repro") # CJ Suggest publishing (at least aminimal working version) of the repro package to CRAN before publishing this paper, to reduce one layer of complexity for users
library("repro")
```

The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with the suggested tools for reproducibility.
For example, the package can check if your computer has all the required software packages installed and ready. <!-- CJ Does repro assume that these have already been installed  by hand? -->

@@

If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.<!-- CJ I think you should explain your philosophy that the package provides additional tutorial instructions depending on the user's situation before diving into the actual tutorial. I foresee that people are going to be puzzled when the tutorial steps present them with additional mini tutorials. Explain in the introduction what the problem is (users need to install andfamiliarize themselves with many different software solutions) and how you solved it (you wrote a package that tells users what dependencies they need to install and how -->
Please make sure that all three commands pass before you continue with the tutorial.
This process is only required once per computer and does not need to be repeated for each project.

# Planning Phase <!-- CJ This reminds me of the WORCS phases, which is fine, but it's the only header that does remind me of these phases. You've not introduced a breakdown of the research process into phases anywhere. Moreover, the text below does not really relate to "planning"; it goes back to reasons for reproducibility, and then an explanation of technical tools (Rmarkdown and Git), neither of which relate to planning! -->

@@

Before incorporating changes from someone else, they can be reviewed and adepted.
This review step can also be useful for collaboration, because you can invite a collaborator to take a look at changes you want to do.
It is therefore sometimes useful to create a pull request even though you have permission to push, because it invites your collaborators to comment or change your proposal.

@@

This process is especially useful if you want the feedback of a more senior researcher, because they do not need to know how to use Git but have a convenient web interface to provide comments or changes.

# Analyzing Data

While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to achieve long-term reproducibility are still missing, namely handling internal and external dependencies.

@@

A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README]).
To do that, it is often crucial to understand how different project components relate to each other.
You cannot run an analysis without data; you cannot render a dynamic document without an analysis, and so forth.
Dependency tracking means that you capture the interrelations between the components of your analysis.
The "components" of an analysis are often loosely mapped onto files.
For a simple project, it might be enough to have a single R Markdown file.
However, while it is easy to understand how to reproduce the end result, this approach quickly gets unwieldy.
To better navigate the data analysis, a researcher often splits it across several scripts.
For example, they create a script to load the data and another to run the analysis.
This modularisation facilitates reuse because the loading script may be used by other projects building on the same data, while the analysis script may be employed on different data and so forth.
The downside is that reproducibility is complicated because instead of reproducing everything in one step, multiple steps are required.

Of course, there a several ways to solve this problem.
One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "main" script that combines all steps.
We found Make to be particularly useful for two reasons.
First, it is language agnostic, therefore fusing steps that are carried out in different programming languages.
Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning.

@@

When you cook something, and you have the necessary ingredients already at hand, would you throw it away just to prepare it again?
Unlikely.
Make behaves the same.
If a file already exists, it does not recreate it unnecessarily.
However, if a requirement is newer than its target, make recreates the target to reflect the updated dependency.
Therefore, when you update the data file, the analysis is rerun.
Because Make knows the whole dependency tree, its caching is smarter than many alternatives.

One alternative that is even smarter than Make is the R package [`targets` [@targets2021]](https://books.ropensci.org/targets/)[^targets].
`targets` can infer more about your project than Make, because it analysis the R code you write to build the dependency tree.
For this inference to work well, you have to write everything in R, and you have to write everything strictly functional.

[^targets]: https://books.ropensci.org/targets/

@@

The `repro` package can automatically write Makefiles for simple scenarios.
For this purpose, `repro` utilizes the YAML metadata:

@@

This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a `Makefile` from them:

@@

The generated `Makefile` automatically contains targets for all R Markdown documents.
To, for example, generate the `README.md` from the `README.Rmd` you can run:

@@

It is common practice to have a target all, that encompasses all important files of the project as the first target.
If you do not specify a target, Make defaults to the first target, which is usually `all`.
`repro` reminds you of that and suggests which file you might want to add to the target `all`.
You can generally change the `Makefile` anyway you want, e.g. to add targets which are not supported by `repro`.
After you have done so, you can then run in the terminal:

@@

And it will fully reproduce the project, if you have all required software installed.

@@

Modern data analysis is only feasible because researchers rely on much external software.
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance .
At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Containers directly address such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
There are two major strategies helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they additionally need on top of these pre-made images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
Our image includes Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`

@@

From the `Dockerfile` you can then build an image.
While you can do so manually we prefer to use Make, see [Internal Dependencies].
The `Makefile` build by `repro` does contain a recipe for creating the `Dockerfile`.

@@

The `Makefile` from `repro` does also include a mechanism to send commands conditionally through the Docker container.
This mechanism is triggered by `DOCKER=TRUE` in the make call:

```bash
# this should be run in the terminal
make DOCKER=TRUE
```
The function `automate()` combines `automate_docker()` & `automate_make()`.

You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manual` to install software beyond R packages.
Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.

@@

Docker can be a bit unfriendly, because it requires a complicated mental model and some knowledge about how to install software on Linux.
An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
`renv` isolates all R packages used in a project and records there version as well as the source from where they where installed.
This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
`renv` therefore is already of great use for reproducibility [@vanlissa2020worcs].
However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS (Basic Linear Algebra Subprograms, responsible for carrying out many computations like matrix multiplication needed in the linear regression) system in use and so forth.
Most importantly the R version itself has to match.
For example, @Peikert2019 illustrate how the generation of random numbers across different R versions is a threat to reproducibility.

@@

The open nature of the here proposed workflow does blur the lines between published and unpublished.
However, even when your Git repository is public, you may want to signify important stages of your project, like preregistration, preprint, submission and acceptance to the journal.
These milestones deserve to be archived.
A project is only then properly archived when it is stored at least at two independent storage locations and reproducible from all locations.
Therefore, you should make extra sure that the project at these points is reproducible.
To ensure reproducible, you could ask a coworker or collaborator to reproduce the project independently from you.
The sections [Creating a DOI] and [Creating a Release] describe how you can easily create three different openly accessible longterm archived copies of your project.

When you disseminate the paper across more locations than just GitHub it can be ambiguous which version of code and data have created the document (e.g. the manuscript.pdf).
To create a clear link between the repository version and some end result you can automatically include the hash of the commit.
Either you use Git from within R:

```{r}
system2("git", c("rev-parse",  "--short",  "HEAD"), stdout = TRUE)
```

Or you can use the `gert` package:

@@

The approach we advocate for distributes the research project, e.g. across GitHub, preregistration platforms, the publishers site, and more.
Digital Object Identifiers (DOI) tie all these locations together under one link [@derisiWhatWhysDOIs2003].
Not only can a DOI resolve to several locations, several parts of a research project can get their own DOI besides the DOI of the publication.
This is useful to link or cite a specific part of a research project, such as the preregistration, an illustration in the paper or the data that was gathered.
The DOI can also contain information about the license of the object it refers to.
<!-- public releases like preregistration, preprint -->
zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides doi for these archives.
While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
Please note that you can only link public GitHub repositories to zenodo.
You may log into zenodo.org through your GitHub account.

@@

After you have linked a GitHub repository you trigger the archivation by creating a GitHub release, as described in the next section.

@@

Unfortunately, this image can only be used when Docker is installed.
To remove any software dependency including Docker from the archived project, one can convert the Docker image to a full system image.
Such system image can subsequently be booted on any personal computer without relying on any software or service and therefore grantees reproducibility for decades to come.
Currently, `repro` does not facilitate this conversion, but we hope to include this feature soon.
